---
title: "Common Assistive Technologies"  
author: 
  - "Mick McQuaid"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css:
      - xaringan-themer.css
      - "https://cdnjs.cloudflare.com/ajax/libs/animate.css/3.7.0/animate.min.css"
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---
class: animated, fadeIn, middle
layout: true

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_solarized_light(
  text_font_google = google_font("Fira Sans", "300"),
  header_font_google = google_font("Fira Sans Condensed","600"),
  code_font_google = google_font("Fira Code")
)
```

<style type="text/css">

/* .remark-slide-content h1 {
  font-size: 600%;
}
*/

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 4px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: red;
}
.scale-30 img { width: 30%; height: 30%; }
.scale-40 img { width: 40%; height: 40%; }
.scale-45 img { width: 45%; height: 45%; }
.scale-60 img { width: 60%; height: 60%; }
.scale-70 img { width: 70%; height: 70%; }
.scale-75 img { width: 75%; height: 75%; }
.scale-80 img { width: 80%; height: 80%; }
.scale-85 img { width: 85%; height: 85%; }
.big { font-size: 200%; }
.bigger { font-size: 400%; }
.biggest { font-size: 600%; }
.footer {
  position: absolute;
  bottom: 10px;
}
</style>

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
xaringanExtra::use_tile_view()
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = TRUE  #<<
)
```
---
class: title-slide

# Common Assistive Technologies

### types, evolution, mainstream technologies, getting used to it

### slideshow by Mick McQuaid

.footer[This is a slideshow. Use the arrow keys to navigate.]

---
class: animated, fadeIn

# This is a test of h1

## This is a test of h2

### This is a test of h3

This is a test of body copy. It should appear as Fira Sans

```r
# This tests code. It should be Fira Code.
mtcars |>
mpg >= 20 |>
  head()
```

---

.biggest[END]

---
class: middle

# Colophon

This slideshow was produced using an R package called *Xaringan*.

The theme is *Solarized Light*.

The fonts are Fira Sans (body), Fira Sans Condensed (headings), and Fira Code.

This slideshow is on GitHub in the `mickmcq/accessibilitySlides` repo.
