class: animated, fadeIn, middle
layout: true

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_solarized_light(
  text_font_google = google_font("Fira Sans", "300"),
  header_font_google = google_font("Fira Sans Condensed","600"),
  code_font_google = google_font("Fira Code")
)
```

```{r, load_refs, include=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "authoryear",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("index.bib")
```

<style type="text/css">

/* .remark-slide-content h1 {
  font-size: 600%;
}
*/

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 4px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: red;
}
.scale-30 img { width: 30%; height: 30%; }
.scale-40 img { width: 40%; height: 40%; }
.scale-45 img { width: 45%; height: 45%; }
.scale-60 img { width: 60%; height: 60%; }
.scale-70 img { width: 70%; height: 70%; }
.scale-75 img { width: 75%; height: 75%; }
.scale-80 img { width: 80%; height: 80%; }
.scale-85 img { width: 85%; height: 85%; }
.scale-125 img { width: 125%; height: 125%; }
.big { font-size: 200%; }
.bigger { font-size: 400%; }
.biggest { font-size: 600%; }
.footer {
  position: absolute;
  bottom: 10px;
}
</style>

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
xaringanExtra::use_tile_view()
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         #<<
  mute_unhighlighted_code = TRUE  #<<
)
```
---
class: title-slide


