---
title: "More R"
format: html
---

```{r}
pacman::p_load(tidyverse)
```

# Homework

```{r}
x <- c(0,3.7,0,3.7,3.9,3.8,3.8,3.8,3.8,3.7,3.8)
stem(x)
```

# Demo with more data sets

## Fatal Police Shootings

```{r}
load("fatal_police_shootings.rda")
ls()
str(fatal_police_shootings)
df <- fatal_police_shootings
df$date <- ymd(df$date)
str(df)
summary(df$date)
# stem(df$date)
stem(as.integer(df$date))
```

```{r}
str(df)
df$fleeing <- ifelse(df$flee=="Not fleeing",0,1)
m <- glm(fleeing ~ age,data=df,family="binomial")
summary(m)
```

```{r}
table(df$race)
df |> ggplot(aes(race,age)) +
  geom_boxplot() +
  geom_jitter(alpha=0.1)
```

```{r}
df |> count(armed) |> arrange(desc(n))
```

Is there a difference between threat levels if a body camera is turned on?

```{r}
tbl<-table(df$threat_level,df$body_camera)
chisq.test(tbl)
```

## Sleep in mammals

```{r}
load("mammals.rda")
ls()
str(mammals)
df <- mammals
m <- lm(total_sleep ~ brain_wt,data=df)
summary(m)
```

```{r}
df |> select(total_sleep,species) |> arrange(total_sleep) |> print(n=62)
```

```{r}
pacman::p_load(GGally)
df |> select(-species) |> ggpairs()
df |> select(-species,-gestation,-exposure,-danger,-body_wt,-predation) |> ggpairs()
pacman::p_load(corrgram)
df |> select(-species,-gestation,-exposure,-danger,-body_wt,-predation) |> corrgram(upper.panel=panel.pie)
```

There are loads of other options for scatterplot matrices, known in the R Graph Gallery as [correlograms](https://r-graph-gallery.com/correlogram.html).

```{r}
pacman::p_load(ellipse)
pacman::p_load(RColorBrewer)
data <- cor(df[,2:7])
my_colors <- brewer.pal(5,"Spectral")
my_colors <- colorRampPalette(my_colors)(100)
ord <- order(data[1,])
data_ord <- data[ord,ord]
plotcorr(data_ord, col=my_colors[data_ord*50+50], mar=c(1,1,1,1) )
```

This didn't work well, so my first guess is that I misspelled something. To check that, I'll use the above code to reproduce the example in the R Graph Gallery.

```{r}
data <- cor(mtcars)
my_colors <- brewer.pal(5,"Spectral")
my_colors <- colorRampPalette(my_colors)(100)
ord <- order(data[1,])
data_ord <- data[ord,ord]
plotcorr(data_ord, col=my_colors[data_ord*50+50], mar=c(1,1,1,1) )
```

This plots correctly, so there must be something else wrong. The next step is to find out what `ord` and `data_ord` are.

```{r}
ord
data_ord
```

Now I see the problem! Everything but but `brain_wt` and `body_wt` has NA values (three such values, I think).

```{r}
# help(cor)
data <- cor(df[,2:7],use="complete.obs")
my_colors <- brewer.pal(5,"Spectral")
my_colors <- colorRampPalette(my_colors)(100)
ord <- order(data[1,])
data_ord <- data[ord,ord]
plotcorr(data_ord, col=my_colors[data_ord*50+50], mar=c(1,1,1,1) )
```

## Employment Resumes

```{r}
load("resume.rda")
ls()
str(resume)
df <- resume
m <- glm(received_callback ~ .,data=df,family="binomial")
summary(m)
m <- glm(received_callback ~ race,data=df,family="binomial")
summary(m)
m <- glm(received_callback ~ gender,data=df,family="binomial")
summary(m)
(tbl <- table(df$gender,df$race))
chisq.test(tbl)
(tbl <- table(df$received_callback,df$race))
chisq.test(tbl)
```

## Sleep Deprivation

```{r}
load("sleep_deprivation.rda")
ls()
str(sleep_deprivation)
df <- sleep_deprivation
table(df$sleep,df$profession)
m <- glm(profession ~ sleep,data=df,family="binomial")
summary(m)
m <- glm(sleep ~ profession,data=df,family="binomial")
summary(m)
df$sleep <- ordered(df$sleep, levels = c("<6","6-8",">8"))
m <- glm(sleep ~ profession,data=df,family="binomial")
summary(m)
m <- glm(profession ~ sleep,data=df,family="binomial")
summary(m)
```

Oops! It is only by coincidence that we get an output for each of these models. The `glm()` function is not applicable to multinomial models unless we do a lot of math on it. Even then, it's not universally applicable. Much better is to actually construct a multinomial logistic regression model. We did that last time with the `multinom()` function, but the output was hard to interpret. Let's use the `tidy()` function to make it easier to read.

```{r}
pacman::p_load(nnet)
pacman::p_load(broom)
pacman::p_load(kableExtra)
m <- multinom(sleep ~ profession,data=df)
tidy(m, conf.int = TRUE) |> kable() |> kable_styling("basic",full_width=FALSE)
```

The `kable` stuff is because, by default, the tibble output format of `tidy()` only shows the first ten letters of a `chr` column, which is not enough to distinguish between them, given that they are all prefixed with the word "profession." The `kableExtra` package provides the tools to construct a better-looking html table.

```{r}
df$sleep <- ordered(df$sleep, levels = c(">8","6-8","<6"))
m <- multinom(sleep ~ profession,data=df)
tidy(m, conf.int = TRUE) |> kable() |> kable_styling("basic",full_width=FALSE)
```

## Murders (not recommended for analysis)

```{r}
load("murders.rda")
ls()
str(murders)
df <- murders
m <- lm(annual_murders_per_mil ~ .,data=df)
summary(m)
```

When people are unemployed, their thoughts turn to murder. Wait, wut?

## NBA players 2018--2019

```{r}
load("nba_players_19.rda")
ls()
str(nba_players_19)
df <- nba_players_19
df |> ggplot(aes(position,height)) +
  geom_boxplot() +
  geom_jitter(alpha=0.2)
```

## Poker (anonymous)

```{r}
load("poker.rda")
ls()
str(poker)
df <- poker
stem(df$winnings)
df |> ggplot(aes(winnings)) +
  geom_histogram()
```

## SAT and GPA data

```{r}
load("satgpa.rda")
ls()
str(satgpa)
df <- satgpa
# m <- glm(sex ~ .,data=df,family="binomial")
table(df$sex)
df$sex <- ifelse(df$sex == 2,0,1)
m <- glm(sex ~ .,data=df,family="binomial")
summary(m)
```

Who knows which sex is which? Anyway.

```{r}
m <- lm(sat_m ~ hs_gpa,data=df)
summary(m)
m <- lm(fy_gpa ~ sat_m,data=df)
summary(m)
```
